---
title: "Final Project"
output: html_document
date: "2023-12-17"
---
For more detailed report, please follow the link to see the complete abstract. [https://docs.google.com/document/d/1sm57L_k4g6a9OZ5qYjTjwThpYfCUf6fh_lsfIz1AdBs/edit?usp=sharing]

# Import Dataset
```{r}
#import dataset
library(haven)
library(purrr)
library(dplyr)
library(ggplot2)
library(tidyr)
trial_path = file.path("PDS_DSA_20050203")
trial_files = list.files(trial_path, pattern = "*_pds2019.sas7bdat")
dl = map(file.path(trial_path, trial_files), ~ read_sas(.x))
names(dl) = gsub("*_pds2019.sas7bdat", "", trial_files)
```
# Get merged table and get the target variables into a new table
```{r}
adrsp = dl$adrsp
adrsp_trt = left_join(adrsp, dl$adsl, by = "SUBJID")
adrsp_trt$arm = adrsp_trt$ATRT

#get the target columns
selected_data <- adrsp_trt|>
  select(SUBJID, DTH, SEX, DIAGTYPE,B_WEIGHT,arm)|>
  distinct()
```

#Get Data Summary Table using function in our package
```{r}
library(bis620.2023)
summarizeData(selected_data)
```

# Data Exploration
```{r}
#SEX
demo <- selected_data
demo_counts= demo |>
  group_by(DTH, SEX)|>
  summarize(count = n(), .groups = "drop")

ggplot(demo_counts, aes(x = DTH, y = count)) +
  geom_col() + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~ SEX)

```

```{r}
#DIAGTYPE
demo <- selected_data
demo_counts= demo |>
  group_by(DTH, DIAGTYPE)|>
  summarize(count = n(), .groups = "drop")

ggplot(demo_counts, aes(x = DTH, y = count)) +
  geom_col() + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~ DIAGTYPE)
```
```{r}
#B_WEIGHT
boxplot(selected_data$B_WEIGHT, main = "boxplot of Baseline Weight")
```

```{r}
#arm
demo <- selected_data
demo_counts= demo |>
  group_by(DTH, arm)|>
  summarize(count = n(), .groups = "drop")

ggplot(demo_counts, aes(x = DTH, y = count)) +
  geom_col() + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~ arm)

```


# Filter the table using a function in our package
```{r}
#filter_by_sex
library(bis620.2023)
#male
male_data <- filterTrialsByCategory_of_X(selected_data, "SEX", "Male")
#female
female_data <- filterTrialsByCategory_of_X(selected_data, "SEX", "Female")
```

# Divide the data based on the B-WEIGHT into two groups -- < mean & > mean using a function in our package
```{r}
#male
m_dividing_criteria <- mean(male_data$B_WEIGHT)
#Below the average group
male_below_average <- filter_by_specifiedvalue(male_data, "B_WEIGHT", m_dividing_criteria)
#Above the average group
male_over_average <- anti_join(male_data, male_below_average, by = "SUBJID")

#female
f_dividing_criteria <- mean(female_data$B_WEIGHT)
#Below the average group
female_below_average <- filter_by_specifiedvalue(female_data, "B_WEIGHT", f_dividing_criteria)
#Above the average group
female_over_average <- anti_join(female_data, female_below_average, by = "SUBJID")

```

# Model Development
1. male_below_average Group
```{r}
#male_below_average Group
library(dplyr)
library(stats)

df <- male_below_average %>%
  mutate(DIAGTYPE = as.factor(DIAGTYPE),
         arm = as.factor(arm))
#Since the sex is fixed, Fit the logistic regression model without the SEX variable
model1 <- glm(DTH ~ DIAGTYPE + B_WEIGHT + arm, family = binomial(link = "logit"), data = df)

summary(model1)
```
2. male_over_average Group
```{r}
#male_over_average Group
library(dplyr)
library(stats)

df <- male_over_average %>%
  mutate(DIAGTYPE = as.factor(DIAGTYPE),
         arm = as.factor(arm))
#Since the sex is fixed, Fit the logistic regression model without the SEX variable
model2 <- glm(DTH ~ DIAGTYPE + B_WEIGHT + arm, family = binomial(link = "logit"), data = df)

summary(model2)
```

3. female_below_average Group
```{r}
#male_over_average Group
library(dplyr)
library(stats)

df <- female_below_average %>%
  mutate(DIAGTYPE = as.factor(DIAGTYPE),
         arm = as.factor(arm))
#Since the sex is fixed, Fit the logistic regression model without the SEX variable
model3 <- glm(DTH ~ DIAGTYPE + B_WEIGHT + arm, family = binomial(link = "logit"), data = df)

summary(model3)
```
4. female_over_average Group
```{r}
#male_over_average Group
library(dplyr)
library(stats)

df <- female_over_average %>%
  mutate(DIAGTYPE = as.factor(DIAGTYPE),
         arm = as.factor(arm))
#Since the sex is fixed, Fit the logistic regression model without the SEX variable
model4 <- glm(DTH ~ DIAGTYPE + B_WEIGHT + arm, family = binomial(link = "logit"), data = df)

summary(model4)
```



